### 2018-04-28 doc/paper.tex
Filling out the notation section.
This is (for now) just a copy of the referenced sections.

\usepackage{amsthm,amssymb,amsmath}
%% H/T: https://tex.stackexchange.com/a/202168/32270
\usepackage{textcomp}
%% H/T: https://tex.stackexchange.com/a/56877/32270
\usepackage{algorithm}
\usepackage{algpseudocode}
%% H/T: https://tex.stackexchange.com/a/244805
\usepackage{booktabs,siunitx}

@@

To address ill-conditioned problems, error-free transformations can
be applied in \textit{compensated algorithms} to account for roundoff.
Error-free transformations were studied in great detail in \cite{Ogita2005}
and open a large number of applications.
In \cite{langlois_et_al:DSP:2006:442}, a compensated Horner's algorithm was
described to evaluate a polynomial in the monomial basis. In \cite{Jiang2010},
a similar method was described to perform a compensated version of the de
Casteljau algorithm. In both cases, the \(\cond{p, s}\) factor is moved
from \(\mach\) to \(\mach^2\) and the computed value is as accurate
as if the computations were done in twice the working precision. For example,
the compensated de Casteljau algorithm (\texttt{CompDeCasteljau}) satisfies
\begin{equation}\label{de-casteljau-2-error}
  \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \mathcal{O}\left(\mach^2\right).
\end{equation}
For problems with \(\cond{p, s} < 1 / \mach^2\), the relative error
is \(\mach\), i.e. accurate to full precision, aside from rounding to the
nearest floating point number.

@@

The paper is organized as follows. Section~\ref{notation} introduces some
basic notations and results about floating point arithmetic, error-free
transformations and the de Casteljau algorithm. In Section~\ref{compensated-2},
the compensated algorithm for polynomial evaluation from \cite{Jiang2010} is
reviewed and notation is established for the expansion. In
Section~\ref{compensated-k}, the \(K\) compensated algorithm is provided and
a forward error analysis is performed. Finally, in Section~\ref{numerical} we
give numerical tests to illustrate the practical efficiency of our algorithms.

@@

>>>>

\usepackage{amsthm,amssymb,amsmath}
%% H/T: https://tex.stackexchange.com/a/202168/32270
\usepackage{textcomp}
%% H/T: https://tex.stackexchange.com/a/56877/32270
\usepackage{algorithm}
\usepackage{algpseudocode}
%% H/T: https://tex.stackexchange.com/a/244805
\usepackage{booktabs,siunitx}
%% H/T: https://tex.stackexchange.com/a/4217/32270
\usepackage{mathtools}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}

@@

To address ill-conditioned problems, error-free transformations (EFT) can
be applied in \textit{compensated algorithms} to account for roundoff.
Error-free transformations were studied in great detail in \cite{Ogita2005}
and open a large number of applications.
In \cite{langlois_et_al:DSP:2006:442}, a compensated Horner's algorithm was
described to evaluate a polynomial in the monomial basis. In \cite{Jiang2010},
a similar method was described to perform a compensated version of the de
Casteljau algorithm. In both cases, the \(\cond{p, s}\) factor is moved
from \(\mach\) to \(\mach^2\) and the computed value is as accurate
as if the computations were done in twice the working precision. For example,
the compensated de Casteljau algorithm (\texttt{CompDeCasteljau}) satisfies
\begin{equation}\label{de-casteljau-2-error}
  \frac{\left|p(s) - \mathtt{CompDeCasteljau}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \mathcal{O}\left(\mach^2\right).
\end{equation}
For problems with \(\cond{p, s} < 1 / \mach^2\), the relative error
is \(\mach\), i.e. accurate to full precision, aside from rounding to the
nearest floating point number.

@@

The paper is organized as follows. Section~\ref{notation} introduces some
basic notation and results about floating point arithmetic, error-free
transformations and the de Casteljau algorithm. In Section~\ref{compensated-2},
the compensated algorithm for polynomial evaluation from \cite{Jiang2010} is
reviewed and notation is established for the expansion. In
Section~\ref{compensated-k}, the \(K\) compensated algorithm is provided and
a forward error analysis is performed. Finally, in Section~\ref{numerical} we
give numerical tests to illustrate the practical efficiency of our algorithms.

@@

Throughout this paper we assume that the computation in floating point
arithmetic obeys the model
\begin{equation}
  a \star b = \fl{a \circ b} = (a \circ b)(1 + \eps_1) =
  (a \circ b) / (1 + \eps_2)
\end{equation}
where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
\in \left\{+, -, \times, \div\right\}\) and \(\left|\eps_1\right|,
\left|\eps_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
and \(\star\) represents the floating point computation, e.g.
\(a \oplus b = \fl{a + b}\). We also assume that the computed result of
\(\alpha \in \mathbf{R}\) in floating point arithmetic is denoted by
\(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
all floating point numbers (see \cite{Higham2002} for more details).
Following \cite{Higham2002}, we will use the following classic properties in
error analysis.

\begin{enumerate}
  \item If \(\delta_i \leq \mach\), \(\rho_i = \pm 1\), then
      \(\prod_{i = 1}^n (1 + \delta_i)^{\rho_i} = 1 + \theta_n\),
  \item \(\left|\theta_n\right| \leq \gamma_n \coloneqq
      n \mach / (1 - n \mach)\),
  \item \((1 + \theta_k)(1 + \theta_j) \leq (1 + \theta_{k + j})\),
  \item \(\gamma_k + \gamma_j + \gamma_k \gamma_j \leq \gamma_{k + j}\),
  \item \((1 + \mach)^j \leq 1 / (1 - j \mach)\).
\end{enumerate}

Now let us introduce some results concerning error-free transformation (EFT).
For a pair floating point numbers \(a, b \in \mathbf{F}\), there exists
a floating point number \(y\) satisfying \(a \circ b = x + y\) where
\(x = \fl{a \circ b}\) and \(\circ \in \left\{+, -, \times\right\}\).
The transformation \((a, b) \longrightarrow (x, y)\) is regarded as an
error-free transformation. The error-free transformation algorithms of the
sum and product of two floating point numbers used later in this paper are
the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
algorithm by Dekker \cite{Dekker1971}, respectively. The following theorem
exhibits the important properties of \texttt{TwoSum} and \texttt{TwoProd}
algorithms (see \cite{Ogita2005}).

\begin{theorem}[\cite{Ogita2005}]
For \(a, b \in \mathbf{F}\) and \(x, y \in \mathbf{F}\), \texttt{TwoSum} and
\texttt{TwoProd} satisfy
\begin{itemize}
  \item \(\left[x, y\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
      \(x + y = a + b\), \(y \leq \mach \left|x\right|\),
      \(y \leq \mach \left|a + b\right|\)
  \item \(\left[x, y\right] = \mathtt{TwoProd}(a, b)\),
      \(x = \fl{a \times b}\), \(x + y = a \times b\),
      \(y \leq \mach \left|x\right|\),
      \(y \leq \mach \left|a \times b\right|\).
\end{itemize}
\end{theorem}

Next, we recall the de Casteljau algorithm:

\begin{theorem}[\cite{Mainar1999}]
If \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) and \(\mathtt{DeCasteljau}(p, s)\)
is the value computed by the de Casteljau algorithm then
\begin{equation}
\left|p(s) - \mathtt{DeCasteljau}(p, s)\right| \leq \gamma_{2n}
\sum_{j = 0}^n \left|b_j\right| B_{j, n}(s).
\end{equation}
\end{theorem}

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}\left(s\right)}{
  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
  \left|p(s)\right|},
\end{equation}
where \(B_{j, n} \geq 0\).
